---
title: "Paper 1: Trunk velocity-dependent Light Touch reduces postural sway during standing"
author: Lee Jones \- Senior Biostatistician \- Statistical Review
date: '`r format(Sys.Date(), "%B %d, %Y")`'
format:
  html:
    embed-resources: true
    theme: journal
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE)
```

```{r}

normalQ<- "QQ-plot looks roughly normal."
notnormalQ<- "QQ-plot indicates the residuals are not normally distributed."
vifprob<-"Overall, when taking into account VIF and SE, the model has collinearity problems."
vifnoprob<-"Overall, when taking into account VIF and SE, the model does not have collinearity issues."
indDesNo<-"The study design is not independent and should be assessed using linear mixed models or generalized estimating equations."
indDesYes<-"The study design seems to be independent."
homYes<-"There is no distinct funnelling pattern observed, supporting homoscedasticity of residuals."
homNo<-"The residuals vs. fitted plot is heteroscedastic with funnelling shape."
somehet<-"Some heteroscedasticity is present in plots, and a sensitivity analysis using weighted or robust regression or wild bootstrapping is recommended."
Cooksnoproblem<-"Cook's Distance and other influence diagnostics indicated no observations of concern."
linyes<-"No linearity violation was observed in either plots or tests."
linno<-"Plots and tests show an unaccounted non-linear relationship may be present."
linplotno<-"Plots show an unaccounted non-linear relationship may be present."
lintestno<-"Tests show an unaccounted non-linear relationship may be present."
linout<-"Linearity issues are likely due to outliers within the data."
p_repro<-"The p-values in this model were reproduced with mean differences for p-values close to zero."
OutliersYes <-"Outliers present with high Cook's distance values indicating a sensitivity analysis should be performed to assess the change in the regression coefficients."
MostCompRep<- "This model was mostly computationally reproducible, with minor rounding errors. P-values were reproduced and had the same interpretation, and regression coefficients did not change direction."
CompRep<-"This model was computationally reproducible, with all reported statistics that were assessed being reproducible."
NotCompRep<-"This model was not computationally reproducible."
 Nosens<-"As this model was not computationally reproducible, inferential reproducibility was not considered, since the original analyses could not be reproduced and therefore, statistical assumptions could not be meaningfully compared or interpreted."

evalAump<-"Evaluate the assumptions of the linear regression models by examining residuals, identifying influential outliers, and assessing multicollinearity among predictors. If any assumptions are violated, address them using appropriate methods."
use_VIF<-"Formally assess multicollinearity using variance inflation factors (VIFs) and examination of coefficient standard errors."
uniqueID <- "Include a unique person identifier (ID) in the dataset to ensure each individual can be distinctly identified."
Scat<-"Consider using Scatterplots to visualise the relationships between variables."
Box<-"Consider using box-plots to visualise the data."
AdjR2<- "Adjusted R^2^ should be reported for multivariable models."
datD<- "Include a data dictionary."
Includedata<- "Include data to reproduce all analyses in the paper and supplementary files."
Fullydescribe<- "Fully describe the model used in the linear regression."
DetailAssump<-"Provide details of linear regression assumption checks."
Code<- "Consider creating a reproducible analysis workflow and sharing the code."
Update_dataAvail<-"Update the data availability statement to reflect the status of the data."
coef95<-"Report linear regression coefficients and 95% confidence intervals."
Analysis<-"Provide tables in the Supporting Information that present all analyses conducted in the paper, including full model outputs such as regression coefficients and all variables used for adjustment."
Missing<-"Provide a clear description of the amount and pattern of missing data, along with the methods used to address missingness in the analyses."
Modelling<-"Ensure that the modelling process is clearly described and applied consistently throughout the paper. The type of model and its key specifications should be included in the footnotes of the tables to improve transparency."
File_format<- "Data underlying statistical analyses should be provided in a standard data format (e.g., CSV, XLSX, or similar) so it is machine readable."
Data_format<-"Provide the data in a recognised long or wide format, with a unique participant identifier included to allow assessment of repeated measures and clustering."
Program<-"Specify the statistical software package, version, and any relevant options, procedures, or settings required to reproduce the analysis."
Stepwise<-"Ensure complete reporting of the stepwise model selection procedure, including the direction of selection (forward, backward, or bidirectional) and the criteria for variable entry and removal."
Multiple_test<-"Acknowledge the issue of multiple testing in the results and/or limitations section, including whether adjustments were applied or why none were considered necessary."
Effect_size<-"Place greater emphasis on the magnitude and practical relevance of regression coefficients, including their confidence intervals, rather than focusing predominantly on statistical significance."
```

```{css}
a { color: blue !important;}
h1 {font-size: 22px !important;}
```

# References

Saini A, Burns D, Emmett D, Song YS (2019) Trunk velocity-dependent Light Touch reduces postural sway during standing. PLoS ONE 14(11): e0224943. <https://doi.org/10.1371/journal.pone.0224943>

Song, Yun Seong, 2019, "Trunk Velocity-Dependent Light Touch Reduces Postural Sway during Standing", <https://doi.org/10.7910/DVN/TB1MGY>, Harvard Dataverse, V1.

# Disclosure

This reproducibility project was conducted to the best of our ability, with careful attention to statistical methods and assumptions. The research team comprises four senior biostatisticians (three of whom are accredited), with 20 to 30 years of experience in statistical modelling and analysis of healthcare data. While statistical assumptions play a crucial role in analysis, their evaluation is inherently subjective, and contextual knowledge can influence judgements about the importance of assumption violations. Differences in interpretation may arise among statisticians and researchers, leading to reasonable disagreements about methodological choices.

Our approach aimed to reproduce published analyses as faithfully as possible, using the details provided in the original papers. We acknowledge that other statisticians may have differing success in reproducing results due to variations in data handling and implicit methodological choices not fully described in publications. However, we maintain that research articles should contain sufficient detail for any qualified statistician to reproduce the analyses independently.

# Methods used in our reproducibility analyses

There were two parts to our study. First, 100 articles published in PLOS ONE were randomly selected from the health domain and sent for post-publication peer review by statisticians. Of these, 95 included linear regression analyses and were therefore assessed for reporting quality. The statisticians evaluated what was reported, including regression coefficients, 95% confidence intervals, and p-values, as well as whether model assumptions were described and how those assumptions were evaluated.

The second part of the study involved reproducing linear regression analyses for papers with available data to assess both computational and inferential reproducibility. Each linear regression model was identified and assigned a number within each paper. A maximum of three linear regression models were selected per paper. For papers with more than three models, the final model or primary models of interest were selected, and any additional models were chosen at random. If no specific models of interest were identified, all three models were selected randomly.

To assess computational reproducibility, differences between the original and reproduced results were evaluated using absolute discrepancies and rounding error thresholds, tailored to the number of decimal places reported in each paper. Results were categorised as *Reproduced*, *Incorrect Rounding*, or *Not Reproduced*, depending on how closely they matched the original values. Each paper was then classified as *Reproduced*, *Mostly Reproduced*, *Partially Reproduced*, or *Not Reproduced*. The mostly reproduced category included cases with minor rounding or typographical errors, whereas partially reproduced indicated substantial errors were observed, but some results were reproduced.

For models deemed at least partially computationally reproducible, inferential reproducibility was further assessed by examining whether statistical assumptions were met and conducting sensitivity analyses, including bootstrapping when appropriate. Changes in standardized regression coefficients, which represent the change in the outcome in standard deviation units per standard deviation increase in a predictor, were assessed for meaningful differences using three thresholds: a relative change of 10% or more, or absolute differences of 0.1 (moderate) and 0.2 (substantial). When non-linear relationships were identified, inferential reproducibility was assessed by comparing model fit measures, including R², Akaike Information Criterion (AIC), and Bayesian Information Criterion (BIC). When the Gaussian distribution was not appropriate for the dependent variable, alternative distributions were considered, and model fit was evaluated using AIC and BIC.

# Summary from statistical review

The main purpose of this paper was to examine a light-touch approach to reduce postural sway. The primary analysis used ANOVA, with repeated-measures (ID) as a fixed effect. While the ANOVA reporting was well described, the linear regression reporting was poor, as only R^2^ was reported. In this case, the linear regression was used to validate the ANOVA findings. The model used is unclear, as the figure suggests a quadratic relationship, but this was not mentioned. The authors mentioned the relationship's direction but reported only R^2^, not the regression coefficients.

# Data availability and software used

The data for the main experiment were provided in a long-formatted Excel spreadsheet, but no accompanying data dictionary was available. Although the authors state that the dataset used for the statistical analyses is available via the Harvard Dataverse, the deposited data supported only ANOVA analyses and did not include the variables required to reproduce the linear regression models. The linear regression analyses were conducted in MATLAB.

# Regression sample

The data to reproduce the linear regression models were not available.

# Computational reproducibility results

The data to reproduce the linear regression models were not available.

# Inferential reproducibility results

The data to reproduce the linear regression models were not available.

# Recommended changes

-   `r Includedata`
-   `r Update_dataAvail`
-   `r Fullydescribe`
-   `r evalAump`
-   `r coef95`

```{r}

repdat <- data.frame(
  paper_ID = as.numeric(1),
  Title = as.character("Trunk velocity-dependent Light Touch reduces postural sway during standing."),
  Summary = as.character("The main purpose of this paper was to examine a light-touch approach to reduce postural sway. The primary analysis used ANOVA, with repeated-measures (ID) as a fixed effect. While the ANOVA reporting was well described, the linear regression reporting was poor, as only R2 was reported. In this case, the linear regression was used to validate the ANOVA findings. The model used is unclear, as the figure suggests a quadratic relationship, but this was not mentioned. The authors mentioned the relationship's direction but reported only R2, not the regression coefficients."),
  Regression_sample = as.character("NA"),
  Data_comments = as.character("The data for the main experiment were provided in a long-formatted Excel spreadsheet, but no accompanying data dictionary was available. Although the authors state that the dataset used for the statistical analyses is available via the Harvard Dataverse, the deposited data supported only ANOVA analyses and did not include the variables required to reproduce the linear regression models."),
  Data_Availability_Statement = as.character("The dataset used for the statistics in this manuscript is now available at Harvard Dataverse, https://doi.org/10.7910/DVN/ TB1MGY."),
  DAS_correct = as.character("No"),
  Data_Notionally_Available = as.character("Yes"),
  Ethics_required_access_data = as.character("NA"),
  Permission_required_access_data = as.character("NA"),
  Data_Available = as.character("No"),
  Anyrawdata = as.character ("Yes"),
  HaveSupInfo = as.character ("No"),
  Data_location = as.character("Harvard Dataverse"),
  Accessible_data_badge  = as.character("No"),
  Data_Reason_not_found  = as.character("Data for ANOVA but no linear regression data identified."),
  Data_Dictionary = as.character("No"),
  Format_file = as.character("Excel"),
  Format_data  = as.character("Long"),
  Package = as.character("MATLAB"),
  Recommended_changes = as.character("Include data so all analyses in the paper can be reproduced. Update the data availability statement to reflect the status of the data. Fully describe the model used in the linear regression. Evaluate the assumptions of the linear regression models by examining residuals, identifying influential outliers, and assessing multicollinearity among predictors. If any assumptions are violated, address them using appropriate methods. If any assumptions are violated, address them using suitable methods. Report linear regression coeffients, 95% confidence intervals."),
  Reproduced = as.character("NA"),
  Rep_comment = as.character("NA"),
  Rating = as.numeric(NA),
  Unclear_N = as.character("NA"),
  Difficult_matching_variables = as.character ("NA"),
  Rounding = as.character ("NA"),
  Unclear_model = as.character ("NA"),
  Typo = as.character ("NA"),
  Data_Cleaning_required = as.character ("NA"),
  Data_Reformatting_required = as.character ("NA"),
  R1 = as.character("NA"),
  R2 = as.character("NA"),
  R3 = as.character("NA"),
  E1 = as.character("NA"),
  E2 = as.character("NA"),
  E3 = as.character("NA"),
  E1_Dist = as.character("NA"),
  E2_Dist  = as.character("NA"),
  E3_Dist  = as.character("NA"),
  E1_Mixed = as.character("NA"),
  E2_Mixed  = as.character("NA"),
  E3_Mixed = as.character("NA"),
  E1_Linkfunction = as.character("NA"),
  E2_Linkfunction = as.character("NA"),
  E3_Linkfunction  = as.character("NA"),
  E1_RSE = as.character("NA"),
  E2_RSE  = as.character("NA"),
  E3_RSE  = as.character("NA"),
  E1_Nonlinear = as.character("NA"),
  E2_Nonlinear  = as.character("NA"),
  E3_Nonlinear  = as.character("NA"),
  E1_Transformation = as.character("NA"),
  E2_Transformation  = as.character("NA"),
  E3_Transformation  = as.character("NA"),
  E1_Regularisation = as.character("NA"),
  E2_Regularisation = as.character("NA"),
  E3_Regularisation  = as.character("NA"),
  E1_boot = as.character("NA"),
  E2_boot  = as.character("NA"),
  E3_boot  = as.character("NA"),
  Comp_reproduce = as.character("NA"),
  Infer_reproduce = as.character("NA"),
  Comp_comment = as.character("NA"),
  Infer_comment = as.character("NA"),
  stringsAsFactors = FALSE
)

write.csv(repdat, file="data/repdat_1.csv")

```

```{r results='hide'}

library(tidyverse)
library(httr)

# capture session information in a text file
sessionInfo() %>% capture.output(file="session_info.txt")

# URL of the file
url <- "https://dataverse.harvard.edu/api/access/datafile/:persistentId?persistentId=doi:10.7910/DVN/TB1MGY/PGWHZK"

# Download the file content
response <- GET(url)

# Check the response status
if (response$status_code == 200) {
  # Create a temporary file to store the downloaded content
  temp_file <- tempfile(fileext = ".tsv")
  
  # Write the content to the temporary file
  writeBin(content(response, "raw"), temp_file)
  
  # Read the tab-delimited file into R
  dat <- read_tsv(temp_file)
  
  print("File downloaded and read successfully!")
  
  # Optionally: display the data
  print(head(data))
} else {
  print("Failed to download the file.")
}



```
